---
title: 'Bass Diffusion Model: Jackery Solar Roof'
author: "Mikayel Harutyunyan"
date: "2026-03-05"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing libraries
```{r}
library(ggplot2)
library(readxl)
library(diffusion)
```

Loading the dataset
```{r}
data <- read_excel("data/cumulative_solar_pv.xlsx", sheet = "Data", col_names = TRUE)

head(data)
```

Data preparation
```{r}
capacity <- data$Capacity_MW

# Converting cumulative capacity to yearly adoption
sales <- diff(capacity)

# Time index
t <- 1:length(sales)

sales
```

STEP 4 - Estimating Bass Model Parameters

Bass model function
```{r}
bass_sales <- function(t, p, q, m) {
  m * ((p + q)^2 / p) * exp(-(p + q) * t) / (1 + (q / p) * exp(-(p + q) * t))^2
}
```

Method 1 - Estimating Bass Parameters using nls()
```{r}
bass_model <- nls(
  sales ~ bass_sales(t, p, q, m),
  start = list(m = max(capacity) * 1.5, p = 0.001, q = 0.2),
  control = nls.control(maxiter = 1000, minFactor = 1/2048, warnOnly = TRUE)
)

summary(bass_model)

# Extracting parameters
p_est <- coef(bass_model)["p"]
q_est <- coef(bass_model)["q"]
m_est <- coef(bass_model)["m"]

p_est
q_est
m_est
```

Method 2 - Estimating Bass Parameters using diffusion library
```{r}
diff_m <- diffusion(sales, type = "bass")

# Extracting and rounding parameters
p_diff <- round(diff_m$w["p"], 4)
q_diff <- round(diff_m$w["q"], 4)
m_diff <- round(diff_m$w["m"], 0)

diff_m
plot(diff_m)
```

Dataframe for comparing the methods
```{r}
# fitted annual sales
fitted_nls <- bass_sales(t, p_est, q_est, m_est)
fitted_diff <- bass_sales(t, p_diff, q_diff, m_diff)

df1 <- data.frame(
  Year = 2000 + t,
  Actual = sales,
  NLS_Fit = fitted_nls,
  DIFF_Fit = fitted_diff
)
```

Annual Sales vs Diffusion 
```{r}
ggplot(df1, aes(x = Year)) +
  geom_point(aes(y = Actual), color = "blue", size = 2) +
  geom_line(aes(y = NLS_Fit), color = "green", linewidth = 1) +
  geom_line(aes(y = DIFF_Fit), color = "orange", linewidth = 1) +
  labs(title = "Annual Solar PV Adoption vs. Bass Model Fits",
       y = "Annual Sales (MW)") +
  theme_minimal()
```


STEP 5 - Predicting Diffusion Path

```{r}
bass_cum <- function(t, p, q, m) {
  m * (1 - exp(-(p + q) * t)) / (1 + (q / p) * exp(-(p + q) * t))
}
```

Forecast for the next 30 years using
```{r}
t_forecast <- 1:35

# Generating predictions
sales_pred_nls <- bass_sales(t_forecast, p_est, q_est, m_est)
sales_pred_diff <- bass_sales(t_forecast, p_diff, q_diff, m_diff)

# Prediction dataframe
df_forecast <- data.frame(
  Year = 2000 + t_forecast,
  NLS_Predicted_Sales = sales_pred_nls,
  DIFF_Predicted_Sales = sales_pred_diff
  
)
```

Plotting historical vs predicted diffusion
```{r}
ggplot() +
  geom_point(data = data.frame(Year = 2000 + t, Actual = sales), 
             aes(x = Year, y = Actual, color = "Historical Sales"), size = 2) +
  geom_line(data = df_forecast, 
            aes(x = Year, y = NLS_Predicted_Sales, color = "Bass NLS Forecast"), linewidth = 1) +
  geom_line(data = df_forecast, 
            aes(x = Year, y = DIFF_Predicted_Sales, color = "Bass DIFF Forecast"), linewidth = 1) +
  scale_color_manual(values = c("Historical Sales" = "blue", "Bass NLS Forecast" = "green", 
                                "Bass DIFF Forecast" = "orange")) +
  labs(title = "Annual Adoption Forecast: Jackery Solar Roof",
       subtitle = "Projection of the Imitation effect through 2035",
       y = "Annual Sales (MW)", x = "Year") +
  theme_minimal()
```

STEP 6 AND 7 - Using Netherlands Residential Solar PV Data

Load the dataset
```{r}
nl_data <- read_excel("data/netherlands_pv.xlsx", sheet = "Data", col_names = TRUE)

head(nl_data)
```

Prepare data for Bass model
```{r}
annual_sales <- diff(nl_data$PV_Systems)
t <- 1:length(annual_sales)
```

Diffusion Method
```{r}
diff_m <- diffusion(annual_sales, type = "bass")

# Extracting and rounding parameters
p_diff <- round(diff_m$w["p"], 4)
q_diff <- round(diff_m$w["q"], 4)
m_diff <- round(diff_m$w["m"], 0)

diff_m
plot(diff_m)
```
Estimate Bass parameters using nls()
```{r}
bass_model <- nls(
  annual_sales ~ bass_sales(t, p, q, m),
  start = list(m = max(nl_data$PV_Systems) * 1.5, p = 0.01, q = 0.3),
  control = nls.control(maxiter = 1000, minFactor = 1/2048, warnOnly = TRUE)
)

summary(bass_model)

# Extracting estimated parameters
p_est <- coef(bass_model)["p"]
q_est <- coef(bass_model)["q"]
m_est <- coef(bass_model)["m"]

p_est
q_est
m_est
```


Forecasting future adoption (2019-2030)
```{r}
t_future <- 1:12
cum_pred <- bass_cum(t_future, p_est, q_est, m_est)
annual_pred <- c(cum_pred[1], diff(cum_pred))
```

Dataframe for plotting
```{r}
df_pred <- data.frame(
  Year = 2019 + t_future - 1,
  Predicted_Annual = annual_pred,
  Predicted_Cumulative = cum_pred
)
```

Add historical data for comparison
```{r}
df_hist <- data.frame(
  Year = nl_data$Year[-1], # removing first year since diff() reduces length
  Actual_Annual = annual_sales,
  Actual_Cumulative = nl_data$PV_Systems[-1]
)
```

Plotting annual adoption: historical vs predicted
```{r}
ggplot() +
  geom_point(data = df_hist, aes(x = Year, y = Actual_Annual), color = "blue", size = 2) +
  geom_line(data = df_pred, aes(x = Year, y = Predicted_Annual), color = "green", linewidth = 1.2) +
  labs(title = "Annual Residential Solar PV Adoption in the Netherlands",
       y = "Annual PV Systems Installed",
       x = "Year") +
  theme_minimal()
```

Plotting cumulative adoption: historical vs predicted
```{r}
ggplot() +
  geom_point(data = df_hist, aes(x = Year, y = Actual_Cumulative), color = "blue", size = 2) +
  geom_line(data = df_pred, aes(x = Year, y = Predicted_Cumulative), color = "green", linewidth = 1.2) +
  labs(title = "Cumulative Residential Solar PV Adoption in the Netherlands",
       y = "Cumulative PV Systems",
       x = "Year") +
  theme_minimal()
```

Computing the time of peak adoption
```{r}
t_peak <- log(q_est / p_est) / (p_est + q_est)
peak_year <- 2019 + t_peak - 1
peak_year
```


FERMI LOGIC

Estimating number of adopters by period using Fermi calculation (m = 3.27 million)
```{r}
m_fermi <- 3272000

cum_pred_fermi <- bass_cum(t_future, p_est, q_est, m_fermi)
annual_pred_fermi <- c(cum_pred_fermi[1], diff(cum_pred_fermi))

df_pred_fermi <- data.frame(
  Year = 2019 + t_future - 1,
  Predicted_Annual = annual_pred_fermi,
  Predicted_Cumulative = cum_pred_fermi
)
```

Plotting cumulative adoption: historical vs predicted
```{r}
ggplot() +
  geom_point(data = df_hist, aes(x = Year, y = Actual_Cumulative), color = "blue", size = 2) +
  geom_line(data = df_pred_fermi, aes(x = Year, y = Predicted_Cumulative), color = "black", linewidth = 1.2) +
  labs(title = "Fermi Cumulative Residential Solar PV Adoption in the Netherlands",
       y = "Cumulative PV Systems",
       x = "Year") +
  theme_minimal()
```